---
title: "Data Science"
output: html_notebook
---

## Importing Libraries

```{r}

require(tidyverse)
require(nycflights13)
require(lubridate)
require(hexbin)

```

## Loading Datasets

```{r}

who

```

### ...
```{r}

who1 <- who %>%
  pivot_longer(new_sp_m014:newrel_f65, names_to = "code", values_to = "cases") #Turning several columns into one (opposite of one hot encoding)

who1

########################################

who2 <- who %>%
  pivot_longer(new_sp_m014:newrel_f65, names_to = "code", values_to = "cases", values_drop_na = TRUE)

who2 

```

### Changing "newrel" to "new_rel" to make the name format consistent
```{r}

who3 <- who2 %>% 
  mutate(code = str_replace(code, "newrel", "new_rel"))

who3

```

### split data into 3 columns based on underscore

```{r}

who4 <- who3 %>% 
  separate(code, c("newold", "type", "sexage"), sep = "_")

who4

```

### Get rid of "newold" because it does not give relevant information, also getting rid of Country names
```{r}

who5 <- who4 %>%
  select(-iso2,-iso3,-newold)

who5

```

### Seperating Sex and Age
```{r}

who6 <- who5 %>%
  separate(sexage, c("sex","age"), sep = 1 )

who6

```

### Pipeline that makes the above all operations
```{r}

tidywho1 <- who %>%
  pivot_longer(new_sp_m014:newrel_f65, names_to = "code", values_to = "cases", values_drop_na = TRUE) %>%
  mutate(code = str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("newold", "type", "sexage"), sep = "_") %>%
  select(-iso2,-iso3,-newold) %>%
  separate(sexage, c("sex","age"), sep = 1 )
  
tidywho1

```

## Relational Data
```{r}

airlines
airports
planes
weather

# How to tell if a column is a key for a specific table? 

```

### ...
```{r}

planes %>% count(tailnum) %>% filter(n>1) #Empty which means that each variable uniquely represents the row

```

### ...
```{r}

weather %>% 
  count(year, month, day, hour, origin) %>% filter(n>1) #Shows how even mistakes happen in the most used datasets

```

### Surrogate key can always be created. Easiest is the row number
```{r}

flights %>%
  mutate(key = row_number()) %>%# Added at the end
  count(key) %>%
  filter(n>1)

flights

```

### ..
```{r}

flights2 <- flights %>%
  select(year:day, hour, origin, dest, carrier) %>%
  left_join(airlines, by = "carrier")

flights2

```

## Creating Data table for different Joins
```{r}

x <- tribble(
  ~key, ~val_x,
  1, "x1",
  2, "x2",
  3, "x3",
)
  
y <- tribble(
  ~key, ~val_y,
  1, "y1",
  2, "y2",
  4, "y3",
)
```

### How joins work
```{r}

# Prioritizes the column that is joining. Left join prio's the left hand table and has NA in the right table (left/right)
# Full join takes all the data values and creates NA in all missing areas. 

```


## Strings
```{r}
## NOT COVERED IN LECTURE, but book talks about it.
```


## Factors
```{r}

x1 <- c("Jan","Apr","Jam", "Mar")


month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

y1 <- factor(x1, levels = month_levels)

y2 <- parse_factor(x1, levels = month_levels) # to show the misspelled error

```

### GSS
```{r}
gss_cat
```

### What are the levels of each Variable
```{r}

gss_cat %>% count(race)
levels(gss_cat$race) #all possible values, NA is possible but does not show up

```

### Bar
```{r}

without <- ggplot(gss_cat, aes(race))+
  geom_bar()

with <- ggplot(gss_cat, aes(race))+
  geom_bar() + 
  scale_x_discrete(drop = FALSE)

with
without

```

### Modifying Factor Order
```{r}

gss_cat

relig <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

relig


```

```{r}

ggplot(relig, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()


```

