---
title: "Data Science: Lecture 4"
output: html_notebook
---

## Exploratory Data Analysis

```{r}

require(tidyverse)
require(nycflights13)
require(lubridate)
require(hexbin)


```


### Catagorical Variables
```{r}

diamonds
# Ordinal data is catagorical but have meaning/order such as the different cuts

```


### Typing Optimizations
```{r}

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x=cut))

diamonds %>%
  ggplot() +
  geom_bar(aes(cut))

```


### Count Optimize
```{r}

diamonds %>% count(cut)

```


## Continous Variables
```{r}

diamonds %>% ggplot() + geom_histogram(aes(x = carat), bins = 30) # 30 bins is the base bin size

diamonds %>% ggplot() + geom_histogram(aes(x = carat), binwidth = .25)

```


```{r}

diamonds %>% count(cut_width(carat, 0.5) %>% fct_rev())

```


## Removing Outlier Values
```{r}

small_d <- diamonds %>%
  filter(carat < 3)

small_d

```


## Distributions
```{r}

ggplot(small_d) + geom_freqpoly(aes(x = carat, color = cut), binwidth = 1/10)
#ggplot(small_d) + geom_freqpoly(aes(x = carat), binwidth = 1/10)
#Frequency poly allows for the inclusion on catagorical variables

```


```{r}

ggplot(small_d) + 
  geom_histogram(aes(x = carat), binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0, 3, 1/10))

```


## 
```{r}

faithful %>% 
  ggplot(aes(x = eruptions)) +
  geom_histogram((binwidth = 0.2))

```

## Removing Outliers

```{r}

d_removed <- diamonds %>% filter(between(y,2,20))

```

## Replace values with N/A
```{r}

d_replaced <- diamonds %>%
    mutate(y = if_else(y<2 | y>20, NA_real_, y))

```

## With Outlier replaced
```{r}

d_replaced %>% filter(is.na(y))

```


## Look at both datasets (one replaced, one removed)
```{r}

ggplot(data = d_replaced) +
  geom_histogram(mapping = aes(x = y), binwidth = .5)
  
ggplot(data = d_removed) +
  geom_histogram(mapping = aes(x = y), binwidth = .5)


# hard to see a difference due to the low amount of outliers

```


## Determine if Depature time is missing
```{r}

cancelled_flts <- flights %>%
    mutate(
      cancelled = is.na(dep_time),
      sched_hour = sched_dep_time %/% 100,
      sched_min = sched_dep_time %% 100,
      sched_dep_time = sched_hour + sched_min/60
)

ggplot(data = cancelled_flts) +
  geom_freqpoly(mapping = aes(x = sched_dep_time, color = cancelled), binwidth = 1/4)
    

```


## Missed section
```{r}

# He uploads the code online if I need it.

```


# Covariation

## Catagorical vs. Continous
```{r}

ggplot(data = diamonds, mapping = aes(x = price)) +
  geom_freqpoly(mapping = aes(color = cut), binwidth = 500) +
  scale_color_hue() # Makes colors easier to see, especially yellow
  

```

## Usefule to find the density
```{r}

ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = price, y = ..density.. , color = cut), binwidth = 500, size = 2) +
  scale_color_hue()

ggplot(data = diamonds, mapping = aes(x = price)) +
  geom_freqpoly(mapping = aes(color = cut), binwidth = 500) +
  scale_color_hue()

```

```{r}

ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price))

# Discussed how box plots are made, and what each seperate part means

ggplot(data = diamonds) +
  geom_violin(mapping = aes(x = cut, y = price), draw_quantiles = .5)

```

## Different mpg per car type
```{r}

ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))

ggplot(data = mpg, mapping = aes(x = fct_reorder(class, hwy, .fun = median()), y = hwy)) +
  geom_boxplot() + # Reorder x axis to the median
  coord_flip() 
  
```

## Contengcy Table
```{r}

diamonds %>% count(color, cut)

```

## 
```{r}

ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))

```

##
```{r}

diamonds %>%
  count(color, cut) %>%
  ggplot() +
    geom_tile(mapping = aes(x = cut, y = color, fill = n))

```


# Continous v.s. Continous

## Missed some stuff

## Hex plot
```{r, fig.width=4, fig.height= 3}

ggplot(data = diamonds) +
  geom_hex(mapping = aes(x = carat, y = price))


```

```{r, fig.width=10, fig.height= 10}

P <- ggplot(data = diamonds) +
  geom_hex(mapping = aes(x = carat, y = price, fill = log2(..count..)))

# Change Color scale
P
P + scale_fill_continuous(low="red", high = "yellow")
P + scale_fill_distiller(pallete = "Y10rRd")
P + scale_fill_distiller(pallete = "Greens")
P + scale_fill_distiller(pallete = "Greys")

```