---
title: "Homework 1"
author: "Preston Robertson"
date: "6/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
library(modelr)
library(dplyr)

```


## Provide a summary of the data. That is, what are the variables. Provide summary statistics for each variable.
```{r}
summary(airquality)
```


```{r}
dat = airquality


dat <- dat %>%
  mutate(dat, ifelse(Month == 5,Actual_Day = Day, 
                     ifelse(Month == 6, Actual_Day = Day + 31, ifelse(Month == 7,Actual_Day = Day + 61, ifelse(Month == 8, Actual_Day = Day + 92, Actual_Day = Day + 123)))))

dat
```



## Make a time series plot for each variable in the airquality data set. Make sure to label each plot carefully. Extra for labeling the months of the data on the x-axis.

```{r pressure, echo=FALSE}

ggplot(airquality, aes(x = Day, y = Ozone)) +
  #geom_ref_line(h = 0, colour = 'red') +
  geom_point() +
  facet_grid(rows = vars(Month))

ggplot(airquality, aes(x = Day, y = Solar.R)) +
  #geom_ref_line(h = 0, colour = 'red') +
  geom_point() +
  facet_grid(rows = vars(Month))

ggplot(airquality, aes(x = Day, y = Wind)) +
  #geom_ref_line(h = 0, colour = 'red') +
  geom_point() +
  facet_grid(rows = vars(Month))

ggplot(airquality, aes(x = Day, y = Temp)) +
  #geom_ref_line(h = 0, colour = 'red') +
  geom_point() +
  facet_grid(rows = vars(Month))

```




## Make side by side boxplots for the ozone data for each month. Do you believe the distribution of the ozone is the same for each month.
```{r}

```




## Compute the correlation and covariance matrix for the variables in the airquality data set.
```{r}

```


